---
title: "Spider sequencing numbers"
author: "David Hemprich-Bennett"
date: "14/06/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
n_spiders <- 396 # This may be slightly wrong: to be checked by Abby after finishing extractions!
```
We have `r n_spiders` samples

```{r, echo = F}
# the number of extracts we NEED to plate,
n_extracts <- n_spiders * 2
```
For our experiment, we want to use 2 primer pairs per spider (ZBJ and ANML). So we need to use `r n_extracts` wells as a minimum

```{r, echo = F}
wells_per_plate <- 96

# the number of plates (to a decimal) that we'll need for n_extracts
n_plates <- n_extracts / wells_per_plate
```

To put all of those extracts in `r wells_per_plate`-well plates, we'll need to use `r n_plates` plates.

```{r, echo = F}
# The number of wells that we're going to need to fill up 
max_wells <- ceiling(n_plates) * wells_per_plate

to_fill <- max_wells - n_extracts
```

However theres no point leaving any empty wells, so we should send them `r ceiling(n_plates)` plates, and fill the `r to_fill` extra wells with blanks, and maybe a few PCR replicates, depending on space.

```{r}
ZBJ_samples <- 4*95
ANML_samples <- 5*95
ANML_replicates <- ANML_samples - n_spiders
```


It looks like we may as well fill `r 4*95` wells with samples for ZBJ, and `r 5*95` wells with ANML. This would mean that we have to include `r (5*95)-n_spiders` PCR replicates of ANML.