---
title: "Spider sequencing numbers"
author: "David Hemprich-Bennett"
date: "24/06/2019"
output: pdf_document
---

```{r setup, include=FALSE}
# Disable code echoing and messaging, as they clutter the document
knitr::opts_chunk$set(echo = F, message = F)

library(tidyverse)
library(lubridate)
library(here)
library(knitr)
```


```{r summary_setup}
# Now lets use the field data

summary_fieldwork <- read.csv(here("data", "raw_data", "field_data", "spider_fielddata.csv"),
  header = T,
  stringsAsFactors = F
)

# Rearrange the summary_fieldwork df nicely for subsequent use
summary_fieldwork <- summary_fieldwork %>%
  # Sort the dates
  mutate(Date = dmy(Date)) %>% 
  mutate(day = day(Date),
         month = month(Date),
         month_name = month(Date, label = T),
         year = year(Date)) %>%
  # replace values with better names: na can be a bad string to have!
  mutate(SiteLampType = gsub("na", "unlit", SiteLampType)) %>%
  # Get rid of the unwanted columns
  select(SpiderCount, SiteName, Treatment, SiteLampType, Date, year, month, 
         month_name, day,Reference, VisitNum) 



  

# make a df which just has the site names and the overall treatments for the locale
visit_treatments <- summary_fieldwork %>%
  select(SiteName, SiteLampType, VisitNum) %>%
  rename(location_type = SiteLampType) %>%
  filter(location_type != 'unlit')

# As the summary fieldwork df doesn't currently provide information on the
# treatment of the unlit sites, add those treatments to summary_fieldwork now 
# using the visit_treatments df we just made
summary_fieldwork <- summary_fieldwork %>%
  left_join(visit_treatments)

```


```{r labdata_load}
# Load in Abby's lab data
lab_data <- read.csv(
  here('data', 'raw_data', 'lab_data', 'DNA Extractions Master.csv'),
  stringsAsFactors = F)
```


```{r combine}
# Combine all the data

combined_data <- lab_data %>%
  rename(Reference = Sample.Collection.ID) %>%
  left_join(summary_fieldwork)

```


---
```{r badextracts}
bad_extracts <- combined_data %>% 
  filter(is.na(combined_data$SiteName)) %>%
  select(Reference, Sample.Unique.ID, BDNA.Conc.ng.uL, Storage.Box,
         Storage.Location)
knitr::kable(bad_extracts, 
             caption = 'The extracts which don\'t match a location in the
             field datasheet')
```

```{r choose_extracts}
to_sequence <- combined_data %>%
  # Get rid of Upper Edington as we it has two lighting types. Also specify
  # that we want to keep control: as it has na values for SiteName and 
  # location_type, it will get filtered out if we don't explicitly keep it
  filter(Sample.Unique.ID == 'CONTROL' | SiteName != "Upper Edington") %>% 
  # also remove LPS, whilst keeping the control
  filter(Sample.Unique.ID == 'CONTROL' | location_type != 'LPS') 

summary_to_seq <- to_sequence %>%
  group_by(location_type, SiteLampType) %>%
  summarise(`Number of extracts` = n())

kable(summary_to_seq, caption = 'Treatments of the spiders to sequence')
```


```{r}
n_spiders <- nrow(to_sequence)
```
We have `r n_spiders` samples, including the extraction control.

```{r}
# the number of extracts we NEED to plate per primer ,
n_extracts <- n_spiders * 2
```
For our experiment we want to use 2 primer pairs per spider (ZBJ and ANML). So we need to use `r n_extracts` wells as a minimum.

```{r}
wells_per_plate <- 95

# the number of plates (to a decimal) that we'll need for each primer pair
n_plates_per_primer <- n_spiders / wells_per_plate

# The number of wells that we're going to need to fill up 
max_wells <- ceiling(n_plates_per_primer) * wells_per_plate

to_fill <- max_wells - n_spiders

```

Things are made more difficult as we shouldn't mix primers on a plate. So we'll have a lot of empty wells to fill out per primer set, as we only need to use `r round(n_plates_per_primer, digits = 2)` plates per primer set, leaving `r to_fill` wells empty. We'll have to use at least `r ceiling(n_plates_per_primer)` plates per primer set, and then fill the empty wells with PCR replicates. 

```{r}
n_ZBJ_plates <- 4
ZBJ_samples <- n_ZBJ_plates * 95
ZBJ_replicates <- ZBJ_samples - n_spiders
n_ANML_plates <- 5
ANML_samples <- n_ANML_plates * 95
ANML_replicates <- ANML_samples - n_spiders

```

It looks like we may as well fill `r n_ZBJ_plates` plates with samples for ZBJ, and `r n_ANML_plates` plates with ANML.
This will give `r ZBJ_samples` wells with samples for ZBJ, and `r 5*95` wells with ANML. This means that we have to include `r ZBJ_replicates` PCR replicates of ZBJ and `r ANML_replicates` PCR replicates of ANML. This will give lots of replicates of ANML, which allows some insight into how effective (or not) PCR replicates are.


```{r}
# Add the number of extracts to be sequenced per spider for each primer pair
conc_order <- order(to_sequence$BDNA.Conc.ng.uL, decreasing = T)
to_sequence$n_ANML <- 1
to_sequence$n_ANML[conc_order[1:ANML_replicates]] <- 2
to_sequence$n_ZBJ <- 1
to_sequence$n_ZBJ[conc_order[1:ZBJ_replicates]] <- 2

# rearrange column orders
to_sequence <- to_sequence %>%
  select(n_ANML, n_ZBJ, everything())

write.csv(to_sequence, here('data', 'to_sequence.csv'),
          row.names = F)

```

```{r}
ANML_duplicates <- to_sequence$Sample.Unique.ID[conc_order[1:ANML_replicates]]
ZBJ_duplicates <- to_sequence$Sample.Unique.ID[conc_order[1:ZBJ_replicates]]

plates <- expand.grid(column = c(1:12),
            row = c('A', 'B', 'C', 'D', 'E', 'F', 'G', 'H'),
            plate = c('ZBJ_01_01', 'ZBJ_01_02', 'ZBJ_01_03', 'ZBJ_01_04',
                      'ANML_01_01', 'ANML_01_02', 'ANML_01_03', 'ANML_01_04', 'ANML_01_05'))

```

Now make the spreadsheet for Guelph:
```{r}
coord_data <- read.csv(here('data', 'raw_data', 'rough_coordinates.csv'), 
                       stringsAsFactors = F)
for_guelph <- to_sequence %>%
  # Add coordinate data
  left_join(coord_data) %>%
  # Retain only the important columns
  select("Sample.Unique.ID", "Date", "Coordinates") %>%
  # Separate the Coordinates into lat/long
  separate(Coordinates, into = c('Lat', 'Long'), sep = ', ') %>%
  # Put the date into a more uniform format
  mutate(Date = ymd(Date)) %>%
  # Add some variables which they want
  mutate(Collectors = 'Douglas Boyes',
         Country = 'United Kingdom',
         Kingdom = 'Animalia',
         Phylum = 'Arthropoda',
         Class = 'Arachnida',
         Order = 'Araneae')

write.csv(for_guelph, here('data', 'sequencing_planning', 'BOLD_fields.csv'),
          row.names = F)
```

