---
title: "Spider sample and cost scenarios"
author: "Dave Hemprich-Bennett"
output: pdf_document
---

```{r setup, include=FALSE}
#### Header ####
## Project:
## Script purpose:
## Date:
## Author: Dave Hemprich-Bennett (hemprich.bennett@gmail.com)
## Notes
##################################################
library(tidyverse)
library(lubridate)
library(ggbeeswarm)
library(here)
library(knitr)
fieldwork_data <- read.csv(here("data", "raw_data", "field_data", "spider_fielddata.csv"),
  header = T,
  stringsAsFactors = F
)

fieldwork_data$Date <- dmy(fieldwork_data$Date)

fieldwork_data <- cbind(fieldwork_data, year = year(fieldwork_data$Date),
                        month = month(fieldwork_data$Date),
                        day = day(fieldwork_data$Date))

fieldwork_data$Bag <- toupper(fieldwork_data$Bag)

fieldwork_data <- fieldwork_data %>%
  unite(sampling_id, Bag, Reference, sep = ' ') %>%
  select(SpiderCount, everything())

table(fieldwork_data$Treatment, fieldwork_data$SiteLampType) 
# number of sampling events

morpho_data <- read.csv(here('data', "raw_data", "field_data", 'spider_morphos.csv'), 
                        stringsAsFactors = F)


morpho_counts <- morpho_data %>%
  group_by(Bag, Pseudospecies, Site, Lighting) %>%
  summarise(n = n()) %>% spread(key = Pseudospecies, value = n) %>%
  unite(sampling_id, Bag, Site, Lighting, sep = ' ')


uber_data <- left_join(fieldwork_data, morpho_counts, by = 'sampling_id')

stripehead_or_straw_sites <- uber_data %>% 
  filter(Stripehead > 1 | !Straws > 1) %>%
  filter(SiteLampType != 'LPS') %>%
  select(Stripehead, Straws, everything())


poss_desirable <- stripehead_or_straw_sites %>%
  filter(VisitNum %in% c(24, 38))


#fieldwork_data<- fieldwork_data %>% # get rid of LPS data, theres not much
#  filter(SiteLampType %in% c('HPS', "LED", "na"))


# different approach, look at sites which have good numbers for both lighting types

net_choosing_sites <- fieldwork_data %>%
  select(SpiderCount, Site, SiteName, month, SiteLampType, Treatment) %>%
  group_by(SiteName, month, SiteLampType) %>%
  summarise(total = sum(SpiderCount)) %>%
  spread(key = month, value = total) %>%
  filter(!is.na(`2`)) %>%
  filter(!is.na(`4`))

summarised_treatments <- net_choosing_sites %>%
  group_by(SiteLampType) %>%
  summarise(Feb = sum(`2`), April = sum(`4`))# %>%
  
manually_edited <- read.csv(here('data', "raw_data", "field_data", 'spider_fielddata_DAVE_MANUALEDITS.csv'), stringsAsFactors = F)
### THIS DATA WAS MANUALLY MUNGED IN A HURRY TO SORT THE SITELAMPTYPE AND MAY HAVE TYPOS. TRUST NOTHING

manually_edited$Date <- dmy(manually_edited$Date)

manually_edited <- cbind(manually_edited, year = year(manually_edited$Date),
                        month = month(manually_edited$Date),
                        day = day(manually_edited$Date))

manual <- manually_edited %>%
  select(SpiderCount, Site, SiteName, month, SiteLampType, Treatment) %>%
  unite(Network, SiteLampType, Treatment) %>%
  group_by(SiteName, month, Network) %>%
  summarise(total = sum(SpiderCount)) %>%
  spread(key = month, value = total) %>%
  replace_na(list(`2`= 0, `4` = 0)) %>%
  group_by(Network) %>%
  summarise(Feb = sum(`2`), April = sum(`4`))# %>%

manual <- manual[-grep('na', manual$Network),]

# manual_costs <- manual %>%
#   mutate(Feb = Feb *10) %>%
#   mutate(April = April *10)
# 
colSums(manual[,2:3])* 10

no_LPS <- manual[grep('HPS|LED', manual$Network),] # Only HPS and LED
colSums(no_LPS[,2:3]) * 10

only_lit <- manual[grep('_Lit', manual$Network),] # Only lit
sum(colSums(only_lit[,2:3])) * 10
#number of samples meeting above criteria


no_LPS_lit <- no_LPS[grep('Lit', no_LPS$Network),]
sum(colSums(no_LPS_lit[,2:3])) * 10

sum(colSums(net_choosing_sites[,3:4]))
```

---

For the `r sum(colSums(manual[,2:3]))` samples if we were to use all collected it would cost roughly £`r sum(colSums(manual[,2:3]))* 10`.  

Breakdown of sample numbers:
```{r, echo = F}
kable(manual)
```

---

For the `r sum(colSums(no_LPS[,2:3]))` samples if we are constrained to using sites with HPS and LED lights, plus their control areas it would cost roughly £`r sum(colSums(no_LPS[,2:3]))* 10`.  

Breakdown of sample numbers:
```{r, echo = F}
kable(no_LPS)
```

---

For the `r sum(colSums(only_lit[,2:3]))` samples only from lit areas it would cost roughly £`r sum(colSums(only_lit[,2:3]))* 10`  

Breakdown of sample numbers:
```{r, echo = F}
kable(only_lit)
```

